Spectroscopic confirmation of very-wide stellar binaries and large-separation comoving pairs from \textsl{Gaia DR1}

### 2017-07-28 paper/comoving-followup.tex
figures and shuffle notebooks

% this paragraph can be simplified
For each source or comparison lamp observation, the two-dimensional image is
bias corrected, flat-field corrected, and trimmed using standard routines
implemented in \package{ccdproc} (\citealt{Craig:2015}).
The 2D spectra are oriented such that the spatial direction is along a row of
pixels, and the dispersion direction is along a column of pixels.
For comparison lamp observations, the central 100 pixel columns are extracted
and summed along the spatial axis, weighted by the inverse-variances of each
pixel.
Source spectral traces were always placed in this region during observations,
and the curvature of the wavelength solution over this region is negligible.
For source observations, 1D spectra are extracted using an empirical
estimate of the line spread function (LSF).
The LSF is assumed to be a Voigt profile with parameters $A$, amplitude; $x_0$,
profile centroid; $\sigma_G$, Gaussian root-variance; $\gamma_L$, Lorentzian
half-width at half maximum, and the background is modeled using a linear
polynomial with parameters $\alpha_1$ and $\alpha_2$, the slope and intercept
at the profile centroid, respectively.
The maximum-likelihood LSF + background model parameters are estimated from
each pixel row using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented in
the \package{scipy} package.
We first determine the LSF width parameters, $\sigma_G$ and $\gamma_L$ from fits
to 16 pixel rows evenly spaced between indices 750 and 850 (i.e. around the
effective center of the dispersion axis).
From these 16 fits, the median profile width parameters, $\sigma_G$ and
$\gamma_L$, are then taken to be the LSF width parameters.
Then, at each of 1600 rows in each source spectrum image, the LSF amplitude and
centroid, and background slope and intercept at line centroid, are then fit
using the same procedure as above but with the width parameters fixed.
The LSF amplitudes are stored as the source fluxes, and the background model
intercepts are stored as the background (sky) fluxes.
Occasionally, nearby sources fall in the slit; these are masked by hand in the
2D images and do not bias the LSF extraction fits.

@@

@@

\figurename~\ref{fig:Halpha-mcmc-corner} shows a \package{corner} plot
(\citealt{Foreman-Mackey:2016}) of all projections of these posterior PDF
samples for a randomly-chosen source (HD 63408, in this case).
The marginal posterior distribution over $x_0$ can be converted to a
distribution over radial velocity; in this case, the estimated (Gaussian
standard deviation) precision is $\approx 2.5~\kms$.
\figurename~\ref{fig:Halpha-mcmc-fit} shows the segment of the source spectrum
around \Ha\ used in the line modeling (black line) with uncertainties (gray
bars).
Top panel shows the linear polynomial + absorption line model computed for the
median sample (dark green line); shaded region shows the 15th to 85th
percentiles computed from all posterior samples.
Bottom panel shows the residuals of the data from the median linear polynomial +
absorption line model (black line) along with the Gaussian process computed from
the median parameters (orange curve) used to model other correlated structure in
the vicinity of \Ha (which can be interpreted as correlated noise).
These figures are produced for all source spectrum fits and visually inspected
to ensure that the MCMC sampling appears to have converged for all cases.

For each source spectrum, with the resulting posterior samples, we compute the
median and median absolute deviation (MAD) of the marginal distribution over
line centroid, $x_0$.
We use the MAD to estimate the standard deviation ($\sigma \approx 1.5\times{\rm
MAD}$, assuming the samples are Gaussian distributed) of the samples and find
that the typical velocity \emph{precision} from measuring the centroid of \Ha\
is between $5$--$10~\kms$.
We report the measured absolute radial velocities along with compiled sky
position, astrometry, and $G$-band magnitude from the \tgas\ catalog and \tmass\
$J$-band magnitudes for all of our comoving star targets in the ``stars'' data
table available with this \documentname; see \tablename~\ref{tbl:data-stars} for
a detailed description of the content of this table.

@@

Several of our targets and all of our calibration targets have previously
measured radial velocities (RVs).
We can therefore estimate the absolute velocity uncertainties by comparing our
derived radial velocities with high-quality prior measurements.
Following previous work (\citealt{Shaya:2011}), we search for previous RV
measurements (in decreasing order of priority) with:
\begin{enumerate}
  \item The Geneva-Copenhagen Survey of the Solar Neighborhood
  (\citealt{Nordstrom:2004}).
  \item The Catalogue of Radial Velocities of Galactic Stars with Astrometric
  Data, the Second Version (\citealt{Kharchenko:2007}).
  \item The General Catalogue of Mean Radial Velocities
  (\citealt{Barbier-Brossat:2000}).
  \item The SIMBAD database (\citealt{Wenger:2000}). We use \package{astroquery}
  (\citealt{Ginsburg:2016}) to retrieve records from the database, which do not
  contain uncertainties but do contain a quality flag. We only retain RVs that
  have a quality flag set to `A' and adopt a $5~\kms$ uncertainty for all such
  values.
\end{enumerate}
% See: paper/figures/Compare with previous observations.ipynb
We find that 155 of the observed stars (mainly calibration targets) have
reliable prior measurements from the literature search described above.
\figurename~\ref{fig:compare-previous} shows the distribution of velocity
differences between our measurements and literature values.
From this distribution (right panel), we compute the median absolute deviation
(MAD) and estimate the standard deviation to be $\approx 20~\kms$, which we
adopt as our global systematic error for all absolute radial velocities.
We note, however, that for computing \emph{relative} velocities for pairs of
nearby stars, it is the RV precision (typically $\approx 5$--$10~\kms$) that
matters.

@@

The inferred 3D velocity differences and values for \llrnew\ and \llrold\ for
all observed candidate comoving pairs are provided with the associated ``pairs''
table; see \tablename~\ref{tbl:data-pairs} for a detailed description of the
content of this table.

@@

Need much larger sample with no cuts on separation or signal-to-noise to study
true separation distribution

Need precise chemical information

But at very large separations, how do you convince yourself that a pair was truly coeval? Same chemistry? To what precision? What about chemical differences from planet formation?

Going forward, will have to use hierarchical mixture models to study population of singles in conjunction

Great reward: Once chemical homogeneity understood amongst stars of same spectral type, can be used to calibrate stellar models across the HR diagram: should have same age, chemistry, etc. Also tests of chemical enrichment due to planetary systems

Abundances - future work, Coronado in prep.! But: throw shade on abundances
because planet engulfment (https://arxiv.org/pdf/1707.02180.pdf). Also Oh et al.
2017b

@@

>>>>

\subsubsection{1D spectrum extraction}

For each source or comparison lamp observation, the two-dimensional image is
bias corrected, flat-field corrected, and trimmed using standard routines
implemented in \package{ccdproc} (\citealt{Craig:2015}).
The 2D spectra are oriented such that the spatial direction is along a row of
pixels, and the dispersion direction is along a column of pixels.
For comparison lamp observations, the central 100 pixel columns are extracted
and summed along the spatial axis, weighted by the inverse-variances of each
pixel.
Source spectral traces were always placed in this region during observations,
and the curvature of the wavelength solution over this region is negligible.
For source observations, 1D spectra are extracted using an empirical
estimate of the line spread function (LSF).
The LSF is assumed to be a Voigt profile with parameters $A$, amplitude; $x_0$,
profile centroid; $\sigma_G$, Gaussian root-variance; $\gamma_L$, Lorentzian
half-width at half maximum, and the background is modeled using a linear
polynomial with parameters $\alpha_1$ and $\alpha_2$, the slope and intercept
at the profile centroid, respectively.
The maximum-likelihood LSF + background model parameters are estimated from
each pixel row using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented in
the \package{scipy} package.
We first determine the LSF width parameters, $\sigma_G$ and $\gamma_L$ from fits
to 16 pixel rows evenly spaced between indices 750 and 850 (i.e. around the
effective center of the dispersion axis).
From these 16 fits, the median profile width parameters, $\sigma_G$ and
$\gamma_L$, are then taken to be the LSF width parameters.
Then, at each of 1600 rows in each source spectrum image, the LSF amplitude and
centroid, and background slope and intercept at line centroid, are then fit
using the same procedure as above but with the width parameters fixed.
The LSF amplitudes are stored as the source fluxes, and the background model
intercepts are stored as the background (sky) fluxes.
Occasionally, nearby sources fall in the slit; these are masked by hand in the
2D images and do not bias the LSF extraction fits.

\subsubsection{Wavelength calibration}

@@

\subsubsection{Radial velocity measurement}

@@

\figurename~\ref{fig:Halpha-mcmc-corner} shows a corner plot of all projections
of these posterior PDF samples for a randomly-chosen source (HD 63408, in this
case).
The marginal posterior distribution over $x_0$ can be converted to a
distribution over radial velocity; in this case, the estimated (Gaussian
standard deviation) precision is $\approx 2.5~\kms$.
\figurename~\ref{fig:Halpha-mcmc-fit} shows the segment of the source spectrum
around \Ha\ used in the line modeling (black line) with uncertainties (gray
bars).
Top panel shows the linear polynomial + absorption line model computed for the
median sample (dark green line); shaded region shows the 15th to 85th
percentiles computed from all posterior samples.
Bottom panel shows the residuals of the data from the median linear polynomial +
absorption line model (black line) along with the Gaussian process computed from
the median parameters (orange curve) used to model other correlated structure in
the vicinity of \Ha\ (which can be interpreted as correlated noise).
These figures are produced for all source spectrum fits and visually inspected
to ensure that the MCMC sampling appears to have converged for all cases.

For each source spectrum, with the resulting posterior samples, we compute the
median and median absolute deviation (MAD) of the marginal distribution over
line centroid, $x_0$.
We use the MAD to estimate the standard deviation ($\sigma \approx 1.5\times{\rm
MAD}$, assuming the samples are Gaussian distributed) of the samples and find
that the typical velocity \emph{precision} from measuring the centroid of \Ha\
is between $5$--$10~\kms$.
We report the measured absolute radial velocities along with compiled sky
position, astrometry, and $G$-band magnitude from the \tgas\ catalog and \tmass\
$J$-band magnitudes for all of our comoving star targets in
\tablename~\ref{tbl:data-stars} available with this \documentname\ (see table
caption for a detailed description of the contents).

@@

Several of our targets and all of our calibration targets have previously
measured radial velocities (RVs).
We can therefore estimate the absolute velocity uncertainties by comparing our
derived radial velocities with high-quality prior measurements.
For calibration sources, we use radial velocity values reported in the source
catalogs (\citealt{Bensby:2014,Pinsonneault:2014}).
For all other sources, following previous work (\citealt{Shaya:2011}), we search
for previous RV measurements (in decreasing order of priority) with:
\begin{enumerate}
  \item The Geneva-Copenhagen Survey of the Solar Neighborhood
  (\citealt{Nordstrom:2004}).
  \item The Catalogue of Radial Velocities of Galactic Stars with Astrometric
  Data, the Second Version (\citealt{Kharchenko:2007}).
  \item The General Catalogue of Mean Radial Velocities
  (\citealt{Barbier-Brossat:2000}).
  \item The SIMBAD database (\citealt{Wenger:2000}). We use \package{astroquery}
  (\citealt{Ginsburg:2016}) to retrieve records from the database, which do not
  contain uncertainties but do contain a quality flag. We only retain RVs that
  have a quality flag set to `A' and adopt a $5~\kms$ uncertainty for all such
  values.
\end{enumerate}
% See: paper/figures/Compare with previous observations.ipynb
We find that 155 of the observed stars (mainly calibration targets) have
reliable prior measurements from the literature search described above.
\figurename~\ref{fig:compare-previous} shows the distribution of velocity
differences between our measurements and literature values.
From this distribution (right panel), we compute the median absolute deviation
(MAD) and estimate the standard deviation to be $\approx 20~\kms$, which we
adopt as our global systematic error for all absolute radial velocities.
We note, however, that for computing \emph{relative} velocities for pairs of
nearby stars, it is the RV precision (typically $\approx 5$--$10~\kms$) that
matters.

@@

The inferred 3D velocity differences and values for \llrnew\ and \llrold\ for
all observed candidate comoving pairs are provided with the associated
\tablename~\ref{tbl:data-pairs} table (see table caption for a detailed
description of the content).

@@

The results presented in this work demonstrate that even at separations larger
than a parsec, a substantial fraction of comoving pair candidates selected from
astrometry alone are confirmed with three-dimensional velocity information.
This is exciting because it confirms the existence of what is likely a
population of nearly-coherent but unbound coeval stars.
It is also encouraging for the prospect of selecting large samples of comoving
pairs purely with astrometric data from future data releases of the \gaia\
mission (radial velocity measurements will only be released for bright stars, $G
< 12$).
However, in this work, we do not claim to report the true number or density of
large-separation comoving pairs.
To model the abundance and formation of comoving pairs, it will be crucial to
select larger samples of pairs without hard cuts on separation, tangential
velocity, signal-to-noise, or likelihood ratio, as were done with the input
sample used in this work.

@@

% Need precise chemical information

% But at very large separations, how do you convince yourself that a pair was truly coeval? Same chemistry? To what precision? What about chemical differences from planet formation?

% Going forward, will have to use hierarchical mixture models to study population of singles in conjunction

% Great reward: Once chemical homogeneity understood amongst stars of same spectral type, can be used to calibrate stellar models across the HR diagram: should have same age, chemistry, etc. Also tests of chemical enrichment due to planetary systems

% Abundances - future work, Coronado in prep.! But: throw shade on abundances
% because planet engulfment (https://arxiv.org/pdf/1707.02180.pdf). Also Oh et al.
% 2017b
